---
title: "Pulling & Binning eMOLT Bottom Temps"
author: "George Maynard"
date: "`r Sys.Date()`"
output: html_document
---

## Function to Download eMOLT data from ERDDAP

```{r emolt_download_source, include=TRUE,message=FALSE,warning=FALSE}
emolt_download=function(
    end_date=lubridate::round_date(
    lubridate::ymd_hms(
      Sys.time()
    ),
    unit="day"
  ),
  days=7){
  ## Download the last week's data from ERDDAP
  ## Seven days before the specified end is the start date for the data
  start_date=end_date-lubridate::days(days)
  ## Use the dates from above to create a URL for grabbing the data
  full_data=read.csv(
    paste0(
      "https://erddap.emolt.net/erddap/tabledap/eMOLT_RT.csvp?tow_id%2Csegment_type%2Ctime%2Clatitude%2Clongitude%2Cdepth%2Ctemperature%2Csensor_type&segment_type=%22Fishing%22&time%3E=",
      lubridate::year(start_date),
      "-",
      lubridate::month(start_date),
      "-",
      lubridate::day(start_date),
      "T00%3A00%3A00Z&time%3C=",
      lubridate::year(end_date),
      "-",
      lubridate::month(end_date),
      "-",
      lubridate::day(end_date),
      "T12%3A02%3A43Z"
    )
  )
  ## Randomly select one point per tow to create a subsampled dataset
  data=full_data[0,]
  tows=unique(full_data$tow_id)
  for(i in tows){
    x=subset(
      full_data,
      full_data$tow_id == i
    )
    data=rbind(
      data,
      x[sample(
        (1:nrow(x)),
        1
      ),]
    )
  }
  return(data)
}
```

## Download the Doppio spatial domain and reproject it

The spatial domain for the Doppio model is in WGS 1984. In order 
to use it to build a grid with cells measurable in km, reproject it to UTM 19S

```{r doppioReproject, include=TRUE,message=FALSE,warning=FALSE}
## Read in the doppio grid
ncin = ncdf4::nc_open("doppio_grid.nc")
## Extract latitudes and longitudes
lat = ncdf4::ncvar_get(
  ncin,
  "lat_rho"
)
lon = ncdf4::ncvar_get(
  ncin,
  "lon_rho"
)

## Create a single column for each
points=data.frame(
  latitude=as.numeric(),
  longitude=as.numeric()
)
for(i in 1:length(lat)){
  x=data.frame(
    latitude=lat[i],
    longitude=lon[i]
  )
  points=rbind(points,x)
  rm(x)
}

## Make it a spatial object
sf.points=sf::st_as_sf(points, coords=c("longitude","latitude"))
sf::st_crs(sf.points)=4326
## Convert it to UTM 19S
sf.points=sf::st_transform(sf.points,32719)
```

## Create a grid from the Doppio spatial extents (10000,10000 is 10 km squares)

```{r gridCreate, include=TRUE,message=FALSE,warning=FALSE}
sf.test=sf::st_make_grid(
  sf.points,
  cellsize=c(42500,42500),
  what="polygons",
  square=TRUE
)
```

## Download the GEBCO bathymetry for plotting over and reproject it

The data downloaded through the marmap package is also in WGS 1984. In order to
use it with our new grid, reproject it into UTM 19S

```{r gebcoReproject, include=TRUE,message=FALSE,warning=FALSE}
bath=marmap::getNOAA.bathy(
  lon1=min(-80.83),
  lon2=max(-56.79),
  lat1=min(35.11),
  lat2=max(46.89),
  resolution=1,
)
## Convert bathymetric object to a raster
sp.bath=marmap::as.raster(bath)
## Reproject raster to UTM 19S
sp.bath=raster::projectRaster(sp.bath,crs=32719)
## Turn the reprojected raster back into a bathymetric object for plotting
sp.bath=marmap::as.bathy(sp.bath)
## Create color ramp
blues=c(
  "lightsteelblue4",
  "lightsteelblue3",
  "lightsteelblue2",
  "lightsteelblue1"
)
```

## Download the eMOLT observations and convert them into a spatial object

```{r eMOLT_download_reproject, include=TRUE,message=FALSE,warning=FALSE}
obs=emolt_download(days=7)
sf.obs=sf::st_as_sf(obs, coords=c("longitude..degrees_east.","latitude..degrees_north."))
## Reproject the spatial object to UTM 19S
sf::st_crs(sf.obs)=4326
sf.obs=sf::st_transform(sf.obs,32719)
sf.obs$cell=NA
```

## Check to see which grid cell an observation falls into

```{r eMOLT_grid_assignment,include=TRUE,message=FALSE,warning=FALSE}
for(i in 1:nrow(obs)){
  sf.obs$cell[i]=sf::st_intersects(sf.obs[i,],sf.test)[[1]]
}
```

## Total observations within a grid cell and assign a color for plotting
```{r gridSummary,include=TRUE,message=FALSE,warning=FALSE}
## Create a new dataframe for frequency of observations
fobs=data.frame(
  cell=unique(sf.obs$cell),
  geometry=NA,
  frequency=NA,
  scalecol=NA
)
for(i in 1:nrow(fobs)){
  x=subset(sf.obs,sf.obs$cell==fobs$cell[i])
  fobs$geometry[i]=x$geometry[1]
  fobs$frequency[i]=nrow(x)
}

nobsval=seq(1,max(fobs$frequency),1)
nobscol=cmocean::cmocean("thermal")(length(nobsval))

for(i in 1:nrow(fobs)){
  fobs$scalecol[i]=nobscol[which(nobsval==fobs$frequency[i])]
}
```

## Plot the output
```{r botGridplotting, include=TRUE,warning=FALSE,message=FALSE}
## Plotting the bathymetry with different colors for land and sea
plot(
  sp.bath,
  step=100,
  deepest.isobath=-1000,
  shallowest.isobath=0,
  col="darkgray",
  image = TRUE, 
  land = TRUE, 
  lwd = 0.1,
  bpal = list(
    c(0, max(sp.bath,na.rm=TRUE), "gray"),
    c(min(sp.bath,na.rm=TRUE),0,blues)
  ),
  xlab="",
  ylab=""
)
for(i in 1:nrow(fobs)){
  plot(sf.test[fobs$cell[i]],col=fobs$scalecol[i],add=TRUE,border='darkgray')
}

legend=round(seq(1,max(nobsval),length.out=10),0)
legend(
  'bottomright',
  legend=legend,
  fill=nobscol[legend],
  title="Observations"
)
```